\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-09 Limber_Comment.tex
edited out Marika (temporarily on her request)
CH typo check read through I think this is ready to send...

\author{Members of the KiDS, CFHTLenS and RCSLenS collaborations; \\
M. Asgari, J. Harnois-D\'eraps, C. Heymans, H. Hildebrandt, S. Joudaki,  M. Kilbinger, F. K\"ohlinger, K. Kuijken, P. Schneider, P. Simon, L. Van Waerbeke, M. Viola}%\altaffilmark{1} ....}
%\email{tbd@tbd}

@@

In this comment paper we discuss the flat-sky and Limber approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation functions. The Limber approximation for cosmic shear was previously explored by \citet{giannantonio/etal:2012} finding that it provided sufficient accuracy for future cosmic shear studies. Recently, however, \citet{kitching/etal:2016} reported that when considering the flat-sky and Limber approximations in concert, tensions between weak lensing and CMB experiments can be reduced. They illustrate the effect with a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.    On initial inspection we found that these claims appeared implausible, for two reasons: First, in a naive estimate, the deviation of flat-sky approximation from a full treatment are expected to be of similar order as the fractional difference between $\theta_{\rm max}$ and $\sin\theta_{\rm max}$,
where $\theta_{\rm max}$ is the maximum angular scale considered in the cosmic shear analysis. In the case of \citep{hildebrandt/etal:2016}, where $\theta_{\rm max}=50$ arcmin, $(\theta_{\rm max} - \sin\theta_{\rm max})/\theta_{\rm max} \approx 3.52 \times 10^{-5}$.  Second, it was shown in \citet[][see their Fig.3]{giannantonio/etal:2012} that the difference between the Limber approximation and exact treatment is far smaller than cosmic variance, even for a full-sky cosmic shear survey.

This comment seeks to explore the content of \citet{kitching/etal:2016} in detail, discussing why we disagree with their findings.  It is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We are unable to replicate key results from \citet{kitching/etal:2016} and show that we do not expect these approximations to impact on the cosmological analysis of current surveys.  In section~\ref{sec:cfhtlens} we present a `one-parameter' cosmological re-analysis of 2D cosmic shear data from \citet{kilbinger/etal:2013} which is inconsistent with the analysis presented in \citet{kitching/etal:2016}.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published and discuss the impact of flat-sky and Limber approximations in our analysis of numerical simulations, CMB lensing and quadratic estimators for cosmic shear power spectrum.

>>>>

\author{Members of the KiDS, CFHTLenS and RCSLenS collaborations; \\
J. Harnois-D\'eraps, C. Heymans, H. Hildebrandt, S. Joudaki,  M. Kilbinger, F. K\"ohlinger, K. Kuijken, P. Schneider, P. Simon, L. Van Waerbeke, M. Viola}%\altaffilmark{1} ....}
%\email{tbd@tbd}
%M. Asgari,

@@

In this comment paper we discuss the flat-sky and Limber approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation functions.  \citet{kitching/etal:2016} report that when considering the flat-sky and Limber approximations in concert, tensions between weak lensing and CMB experiments can be reduced. They illustrate the effect with a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.    On initial inspection we found that these claims appeared implausible, for two reasons: First, in a naive estimate, the deviation of a flat-sky approximation from a full treatment is expected to be of a similar order to the fractional difference between $\theta_{\rm max}$ and $\sin\theta_{\rm max}$,
where $\theta_{\rm max}$ is the maximum angular scale considered in the cosmic shear analysis. In the case of \citet{hildebrandt/etal:2016}, where $\theta_{\rm max}=50$ arcmin, $(\theta_{\rm max} - \sin\theta_{\rm max})/\theta_{\rm max} \approx 3.5 \times 10^{-5}$.  Second, it was shown in \citet[][see their Fig.3]{giannantonio/etal:2012} that the difference between the Limber approximated shear power spectrum and one calculated using an exact treatment is far smaller than cosmic variance, even for a full-sky cosmic shear survey.

This comment seeks to explore the content of \citet{kitching/etal:2016} in detail, discussing why we disagree with their findings.  It is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We are unable to replicate key results from \citet{kitching/etal:2016} and show that we do not expect these approximations to impact on the cosmological analysis of current surveys.  In section~\ref{sec:cfhtlens} we present a `one-parameter' cosmological re-analysis of 2D cosmic shear data from \citet{kilbinger/etal:2013} which is inconsistent with the analysis presented in \citet{kitching/etal:2016}.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published and discuss the impact of flat-sky and Limber approximations in our analysis of numerical simulations, CMB lensing and quadratic estimators for the cosmic shear power spectrum.

### 2016-12-09 cfhtlens.tex
CH typo check read through I think this is ready to send...

\citet{kitching/etal:2016} chose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}. This can be compared to the full five-parameter fit from \citet{kilbinger/etal:2013} who find $S_8=\sigma_8(\Omega_m/0.3)^{0.6} = 0.74 \pm 0.03$.   We argue that this reduction in the relevant error by a factor of $\sim 2$, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} and are unable to confirm the results quoted by \citet{kitching/etal:2016}.   We find the lowest conditional value of $\sigma_8 = 0.746 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.749 \pm 0.023$ for the `Baseline Limber Flat' (LF) case, a maximum change of $\Delta \sigma_8 = 0.0026 = 0.11 \sigma$.  These findings are in disagreement with \citet{kitching/etal:2016} who find a shift of $\Delta \sigma_8 = 0.007 = 0.5\sigma$ between their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) and their `Limber' analysis (which we believe corresponds to the Extended Limber spherical (ELS) case in Table~\ref{tab:Tl_nu}).

@@

We note that \citet{kitching/etal:2016} use their `Flat' approximation for the `Full' non-Limber analysis of the CFHTLenS data.   This is because on a sphere, correlation functions cannot be related to the power spectrum by a Hankel transform (e.g. equation~\ref{eqn:xiGG}) as the power spectrum is not defined for non-integer $\ell$.   In Figure~\ref{fig:Cl_xi}, however, we show that this chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution in the Limber approximated case.  This is not the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.  We would therefore argue that it is more appropriate to use a $T_\ell = 1$ flat-sky approximation when carrying out a `Full' non-Limber analysis of the shear correlation function.

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range of different two-point shear statistics including the shear correlation function, $\xi_\pm$, the aperture-mass dispersion, $\langle M_ {\rm ap} \rangle ^2$ \citep{schneider/etal:1998}, and the COSEBIs, $E_n$ \citep{schneider/etal:2010}.  These statatistics are linearly related to the power spectrum via integrals of the form,
%
\begin{align}
\label{eqn:integ}
& E_n =\int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P(\ell)\;,\\ \nonumber
& \langle M_ {\rm ap} \rangle ^2(\theta)=\int_0^{\infty}\frac{\d \ell \ell}{2\pi} U^2_\theta(\ell) P(\ell)\;,
\end{align}
%
where $W_n(\ell)$ and $U^2_\theta(\ell)$ are defined in \cite{schneider/etal:2010} and \cite{schneider/etal:1998}. The corresponding equations for $\xi_\pm$ is given in \eqref{eqn:xiGG}.
Figure~\ref{fig:filters} shows the integrands of these statistics 
for two cases, normalised to their maximum value. 
The integrands are of the form $\ell F(\ell) P(\ell)$ (see equations~\ref{eqn:xiGG} and \ref{eqn:integ}). For $\xi_\pm$, $P(\ell)$ is equal to the sum of the E and B-mode power spectra. This is one of the main motivators for the development of aperture-mass dispersion. However, $\langle M_ {\rm ap} \rangle ^2$ statistics is a lossy conversion and is biased by small angular separations, where blending of galaxies makes shear measurement challenging/impossible \citep{kilbinger/etal:2006}. As a result, COSEBIs was developed which tackles both shortcomings of $\langle M_ {\rm ap} \rangle ^2$ statistics.
The upper two panels show the integrands\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, $\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and $\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.

@@

\begin{figure}[!htp]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=\textwidth]{figures/IntegAll.pdf} \\
\end{tabular}
\caption{ \small{\label{fig:filters}. Integrand of $\xi_+$ (upper), $\xi_-$ (upper middle), $E_1$ ( lower middle, E-COSEBIs) and $\langle M_{\rm ap} \rangle^2$ (lower panel).
All integrands are of the form $\ell F(\ell) P(\ell)$, where $F(\ell)$ is the corresponding weight-function
for each statistic and $P(\ell)$ is the E-mode convergence power spectrum, with the exception of $\xi_\pm$, for which
$P(\ell)$ is equal to the sum of E and B-mode power spectra. 
Two cases are shown for each statistic as listed in each caption.
For the aperture mass statistic $\theta_{\rm max}=2\theta$ is shown. 
Note that higher order COSEBIs modes generally probe larger $\ell$-modes, 
hence here we only show the lowest mode $E_1$. All values are normalized with respect to their maximum value. }
}
\end{center}
\end{figure}

\subsection{On the Validation of Methods on Simulations}
It is discussed in \citet{kitching/etal:2016} that the validation of cosmic shear estimators on mock catalogues could also be affected by the approximations they investigate.  We agree that a full sky mock should be compared to a full curved sky non-Limber prediction.  It is then suggested, however, that the discrepancies already reported between existing mocks and the predictions could be attributed to these approximations, as opposed to a finite box effect \citep{kiessling/etal:2011, harnois-deraps/etal:2012, harnois-deraps/vanwaerbeke:2015}.  Going beyond flat-sky and Limber could improve the agreement, however this cannot explain the results of \citet{harnois-deraps/vanwaerbeke:2015} who demonstrate that the large-scale under-estimation of power varies with the simulation box size. This volume effect was verified by comparing explicitly the $\xi_+$ statistic at large angular scales between simulation volumes with box sides  ranging from 147$\;\rm{Mpc/h}$ to 515$\;\rm{Mpc/h}$ \citep[see Fig. 5 in][]{harnois-deraps/vanwaerbeke:2015}. The increase of power loss with decreasing box size was shown to be fully-modelled by excluding the super-box $k$-modes in the Limber predictions, something that the approximations discussed in \citet{kitching/etal:2016} would be unable to explain.

@@

\subsection{On the quadratic estimator}

\citet{kitching/etal:2016} state that `the derivative of a Limber-approximated power spectrum is not equal to the Limber-approximation of the derivative'. Based on this statement they conclude that `statistics that use Limber-approximated quadratic estimators to measure the cosmic shear power spectrum are not accurate' and refer explicitly to the quadratic estimator implementation used in \citet{Koehlinger2016} to measure the cosmic shear power spectrum from CFHTLenS. In this quadratic estimator implementation, however, the Limber approximation does not enter. The algorithm was originally presented by \citet{Hu2001} in the context of cosmic shear and has also been applied to other shear data sets (e.g. \citealt{Brown2003, Heymans2005, Lin2012}). Below we briefly summarize this quadratic estimator algorithm and emphasize that the Limber approximation does not enter in there.

The aim of the method is to find the best-fitting power spectra assuming a Gaussian likelihood function $\mathcal{L}$. The likelihood depends on the shear data vector and its covariance $\mathbf{C}$ which consists of the sum of the shear correlation matrix $\mathbf{C}^\mathrm{sig}$ and the noise matrix $\mathbf{C}^\mathrm{noise}$. The shear correlation matrix can be written as a linear combination of piece-wise constant band powers $\mathcal{B}$ approximating the Fourier transforms of the shear correlations, i.e. their power spectra (cf. equations~3,~10 in \citealt{Koehlinger2016}). The best-fitting power spectra are found by iteratively stepping through a Newton--Raphson method in order to find the root of $\diff \mathcal{L}/ \diff \mathcal{B} = 0$ \citep{Bond1998, Seljak1998}. In the subsequent equations of the Newton--Raphson method only derivatives of the full covariance matrix enter, where the derivatives are taken with respect to band powers\footnote{i.e. $\partial \mathbf{C} / \partial \mathcal{B}_A$, where $A$ is a superindex over the type of the band, band-power bin and redshift correlation, see appendix\~A in \citealt{Koehlinger2016}.}.  The only approximation entering in the algorithm is the flat-sky approximation when introducing the Fourier decomposition of the spin-2 shear field.    This approximation is appropriate and accurate for current surveys as $\ell>80$  in all analyses published to date.

>>>>

The \citet{kitching/etal:2016} re-analysis of the 2D CFHTLenS data from \citet{kilbinger/etal:2013} chose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}). This result can be compared to the full five-parameter fit from \citet{kilbinger/etal:2013} who find $S_8=\sigma_8(\Omega_m/0.3)^{0.6} = 0.74 \pm 0.03$.   We argue that this reduction in the relevant error by a factor of $\sim 2$, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} and are unable to confirm the results quoted by \citet{kitching/etal:2016}.   We find the lowest conditional value of $\sigma_8 = 0.746 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.749 \pm 0.023$ for the `Baseline Limber Flat' (LF) case, a maximum change of $\Delta \sigma_8 = 0.0026 = 0.11 \sigma$.  These findings are in disagreement with \citet{kitching/etal:2016} who find a shift of $\Delta \sigma_8 = 0.007 = 0.5\sigma$ between their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) and their `Limber' analysis (which we believe corresponds to the Extended Limber spherical (ELS) case in Table~\ref{tab:Tl_nu}).

@@

We note that \citet{kitching/etal:2016} use their `Flat' approximation for their `Full' non-Limber analysis of the CFHTLenS data.   This is because on a sphere, correlation functions cannot be related to the power spectrum by a Hankel transform (e.g. equation~\ref{eqn:xiGG}) as the power spectrum is not defined for non-integer $\ell$.   In Figure~\ref{fig:Cl_xi}, however, we show that this chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution in the Limber approximated case.  This is not the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.  We would therefore argue that it is more appropriate to use a $T_\ell = 1$ flat-sky approximation when carrying out a `Full' non-Limber analysis of the shear correlation function.

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range of different two-point shear statistics including the shear correlation function, $\xi_\pm$, the aperture-mass dispersion, $\langle M_ {\rm ap} \rangle ^2$ \citep{schneider/etal:1998}, and the COSEBIs, $E_n$ \citep{schneider/etal:2010}.  These statistics are linearly related to the power spectrum via integrals of the form,
%
\begin{align}
\label{eqn:integ}
& E_n =\int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P(\ell)\;,\\ \nonumber
& \langle M_ {\rm ap} \rangle ^2(\theta)=\int_0^{\infty}\frac{\d \ell \ell}{2\pi} U^2_\theta(\ell) P(\ell)\;,
\end{align}
%
where $W_n(\ell)$ and $U^2_\theta(\ell)$ are defined in \cite{schneider/etal:2010} and \cite{schneider/etal:1998}. The corresponding equation for $\xi_\pm$ is given in equation~\ref{eqn:xiGG}.
Figure~\ref{fig:filters} shows the integrands of these statistics 
for two cases, normalised to their maximum value, where the integrands are of the form $\ell F(\ell) P(\ell)$. For $\xi_\pm$, $P(\ell)$ is equal to the sum of the E and B-mode power spectra, motivating the development of the aperture-mass dispersion statistic, $\langle M_ {\rm ap} \rangle ^2$.  This statistic is, however, a lossy conversion and is biased by small angular separations, where blending of galaxies makes shear measurement challenging \citep{kilbinger/etal:2006}.  The COSEBIs statistic tackles both these shortcomings.  The upper two panels of figure~\ref{fig:filters} show the integrands\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, $\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and $\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.

\begin{figure}[!htp]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=\textwidth]{figures/IntegAll.pdf} \\
\end{tabular}
\caption{ \small{\label{fig:filters}. Integrand of $\xi_+$ (upper), $\xi_-$ (upper middle), $E_1$ (lower middle, E-COSEBIs) and $\langle M_{\rm ap} \rangle^2$ (lower panel).
All integrands are of the form $\ell F(\ell) P(\ell)$, where $F(\ell)$ is the corresponding weight-function
for each statistic and $P(\ell)$ is the E-mode convergence power spectrum, with the exception of $\xi_\pm$, for which
$P(\ell)$ is equal to the sum of the E and B-mode power spectra. 
Two cases are shown for each statistic as listed in each caption.
For the aperture mass statistic $\theta_{\rm max}=2\theta$ is shown. 
Note that higher order COSEBIs modes generally probe larger $\ell$-modes, 
hence here we only show the lowest mode $E_1$. All values are normalized with respect to their maximum value. }
}
\end{center}
\end{figure}

@@

\subsection{On the Validation of Methods on Simulations}
It is discussed in \citet{kitching/etal:2016} that the validation of cosmic shear estimators on mock catalogues could also be affected by the approximations they investigate.  We agree that a full sky mock should be compared to a full curved sky non-Limber prediction.  It is then suggested, however, that the discrepancies already reported between existing mocks and the predictions could be attributed to these approximations, as opposed to a finite box effect \citep{kiessling/etal:2011, harnois-deraps/etal:2012, harnois-deraps/vanwaerbeke:2015}.  Going beyond flat-sky and Limber could improve the agreement, however this cannot explain the results of \citet{harnois-deraps/vanwaerbeke:2015} who demonstrate that the large-scale under-estimation of power varies with the simulation box size. This volume effect was verified by comparing the $\xi_+$ statistic at large angular scales between simulation volumes with box sides  ranging from $147\;\rm{Mpc/h}$ to $515\;\rm{Mpc/h}$ \citep[see Fig. 5 in][]{harnois-deraps/vanwaerbeke:2015}. The increase of power loss with decreasing box size was shown to be fully-modelled by excluding the super-box $k$-modes in the Limber predictions, something that the approximations discussed in \citet{kitching/etal:2016} would be unable to explain.

@@

\subsection{On the quadratic estimator of the cosmic shear power spectrum}

\citet{kitching/etal:2016} state that `the derivative of a Limber-approximated power spectrum is not equal to the Limber-approximation of the derivative'. Based on this statement they conclude that `statistics that use Limber-approximated quadratic estimators to measure the cosmic shear power spectrum are not accurate' and refer explicitly to the quadratic estimator implementation used in \citet{Koehlinger2016} to measure the cosmic shear power spectrum from CFHTLenS. In this quadratic estimator implementation, however, the Limber approximation does not enter.

The quadratic estimator algorithm was originally presented by \citet{Hu2001} in the context of cosmic shear and has also been applied to other shear data sets (e.g. \citealt{Brown2003, Heymans2005, Lin2012}).   The aim of the method is to find the best-fitting power spectra assuming a Gaussian likelihood function $\mathcal{L}$. The likelihood depends on the shear data vector and its covariance $\mathbf{C}$ which consists of the sum of the shear correlation matrix $\mathbf{C}^\mathrm{sig}$ and the noise matrix $\mathbf{C}^\mathrm{noise}$. The shear correlation matrix can be written as a linear combination of piece-wise constant band powers $\mathcal{B}$ approximating the Fourier transforms of the shear correlations, i.e. their power spectra (cf. equations~3,~10 in \citealt{Koehlinger2016}). The best-fitting power spectra are found by iteratively stepping through a Newton--Raphson method in order to find the root of $\diff \mathcal{L}/ \diff \mathcal{B} = 0$ \citep{Bond1998, Seljak1998}. In the subsequent equations of the Newton--Raphson method only derivatives of the full covariance matrix enter, where the derivatives are taken with respect to band powers\footnote{i.e. $\partial \mathbf{C} / \partial \mathcal{B}_A$, where $A$ is a superindex over the type of the band, band-power bin and redshift correlation, see appendix\~A in \citealt{Koehlinger2016}.}.  We emphasize that the only approximation entering this algorithm is the flat-sky approximation when introducing the Fourier decomposition of the spin-2 shear field.    This approximation is appropriate and accurate for current surveys as $\ell>80$  in all analyses published to date.

### 2016-12-09 conclusion.tex
CH typo check read through I think this is ready to send...

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat-sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between current weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the important core message of their paper, that future surveys will need to consider these approximations and optimise their statistical analyses accordingly.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{schneider/etal:2010} renders the cosmic shear measurement completely insensitive to the low-$\ell$ scales where these approximations become important.

The treatment of photo-$z$ uncertainties in \cite{kitching/etal:2016} is based on a simple linear shift of redshift distributions. This has been shown to be insufficient to capture the full complexity of the problem.  More importantly though, we suggest that \cite{kitching/etal:2016} have applied the photo-$z$ calibrations from \cite{choi/etal:2016} in the wrong sense. Instead of alleviating the tension between CFHTLenS and Planck a correct application of the offsets would increase the discrepancy, as shown with a recent re-analysis of the CFHTLenS data \citep{joudaki/etal:2016} as well as results from the KiDS project \citep{hildebrandt/etal:2016}.

>>>>

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat-sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between current weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the important core message of their paper, that future surveys will need to consider these approximations and optimise their statistical analyses accordingly.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{schneider/etal:2010} renders the cosmic shear measurement insensitive to the low-$\ell$ scales where these approximations become important.

The treatment of photo-$z$ uncertainties in \cite{kitching/etal:2016} is based on a simple linear shift of redshift distributions. This has been shown to be insufficient to capture the full complexity of the problem.  More importantly though, we suggest that \cite{kitching/etal:2016} have applied the photo-$z$ calibrations from \cite{choi/etal:2016} in the wrong sense. Instead of alleviating the tension between CFHTLenS and Planck, a correct application of the offsets would increase the discrepancy, as shown with a recent re-analysis of the CFHTLenS data \citep{joudaki/etal:2016} as well as results from the KiDS project \citep{hildebrandt/etal:2016}.
